---
title: "ETC3550 Assignment 8"
author: "Chelaka Paranahewa"
output: 
  pdf_document:
    fig_crop: false
---
```{r Setup}
library(fpp3)
```
```{r Data Preparation}
vic_tourism <- tourism %>% filter(State == "Victoria")
# vic_tourism %>%
#     filter(Region == "Lakes", Purpose == "Holiday") %>%
#     gg_tsdisplay(plot_type = 'partial')
```

```{r ARIMA Modelling}
progressr::with_progress(
    fit <- vic_tourism %>%
        model(
            ARIMA = ARIMA(Trips),
            ETS = ETS(Trips)
        )
)
```
```{r}
progressr::with_progress(
    ARIMA_forecasts <- fit %>% select(ARIMA) %>% forecast(h = 16)
)

fit %>% filter(Region == "Melbourne")
ARIMA_forecasts %>%
    filter(Region == "Melbourne") %>%
    autoplot()
# Number of differences is 1 for all non seasonal comps


fit %>% filter(Region == "High Country")
ARIMA_forecasts %>%
    filter(Region == "High Country") %>%
    autoplot()
```

From the Melbourne Region, only the business and visiting graphs had seasonal forecasts. Likely because they were the only purposes that had a seasonal period and also note they had no season differences. While the holiday and other both show case a linear forecast which does have 1 seasonal difference but no seasonal period.

From the High Country Region, only the holiday graph shows a strong seasonal forecast which had 1 seasonal difference and a seasonal period of 4. Visiting has a very subtle seasonality with in the first 4 quarters due to it also having a seasonal period of 4 but it does not have any seasonal differences. 

So from the two regions, it seems that the seasonal period and one difference needs to exist in the ARIMA model to drive the seasonality in the forecast. 