---
title: "ETC3550 Assignment 7"
author: "Chelaka Paranahewa"
output: 
  pdf_document:
    fig_crop: false
---

```{r Setup, echo = FALSE, include = FALSE}
library(fpp3, warn.conflicts = FALSE)
library(broom, warn.conflicts = FALSE)
library(glue, warn.conflicts = FALSE)
# sessionInfo()
```

TODO
Write an R function to generate data from an AR(2) model with parameters to be specified by the user. The first line of your function definition should be

ar2 <- function(n, c, phi1, phi2, sigma)
Here n is the number of observations to simulate, c is the constant, phi1 and phi2 are the AR coefficients and sigma is the standard deviation of the noise.

The noise should be generated using the rnorm() function.

You can set the first two observations equal to c/(1-phi1-phi2). This is the mean of the AR process.

Your function should check that the values of phi1 and phi2 satisfy the stationarity constraints, and if not it should issue a warning. You can use the warning() function for this.

Your function should return a numeric vector of length n containing the generated AR(2) data.
Please submit your solution as a .R file.

```{r Autoregressive Model}
ar2 <- function(n, c, phi1, phi2, sigma) {
    # Input Verifier
    # -1 < phi2 < 1
    if (-1 >= phi2 | phi2 >= 1) warning("Phi2 is not between -1 and 1")

    # phi1 + phi2 < 1
    if (phi1 + phi2 >= 1) warning("Phi1 + Phi2 is greater than or equal to 1")

    # phi2 - phi1 < 1
    if (phi2 - phi1 >= 1) warning("Phi1 - Phi2 is greater than or equal to 1")

    output_vect <- rep(0, n)
    output_vect[1] <- c / (1 - phi1 - phi2)
    output_vect[2] <- c / (1 - phi1 - phi2)

    for (i in 3:n) {
        output_vect[i] <- c +
            phi1 * output_vect[i - 1] +
            phi2 * output_vect[i - 2] +
            rnorm(1, 0, sigma)
    }

    return(output_vect)
}
```


```{r tests}
n <- 400
c <- 3
p1 <- 0
p2 <- -0.9
sig <- 2
ar2(n, c, p1, p2, sig)
```