---
title: "ETC2410 Assignment 2"
author: "Alex Wong, Chelaka Paranahewa, Harjot Channa, Jonas Tiong"

output: 
    pdf_document:
        extra_dependencies: ["mathtools", "amsmath", "flexisym", "breqn"]
        df_print: kable
    latex_engine: xelatex
mainfont: Times New Roman
fontsize: 12pt
---

# **Question 2 (31 Marks)**

```{r Setup, echo = FALSE, warning = FALSE, error = FALSE, message = FALSE}
library(tidyverse)
library(fpp3)
library(knitr)

knitr::opts_chunk$set(
    echo = FALSE,
    warning = FALSE,
    error = FALSE,
    message = FALSE
)

db <- read.csv2(
    file = "Assignment 2/Data/HOUSTNSA.csv",
    sep = ","
)

db <- db %>%
    as_tibble() %>%
    mutate(
        DATE = yearmonth(DATE),
        HOUSTNSA = as.numeric(HOUSTNSA)
    ) %>%
    as_tsibble()

db$Jan <- ifelse(month(db$DATE) == 1, 1, 0)
db$Feb <- ifelse(month(db$DATE) == 2, 1, 0)
db$Mar <- ifelse(month(db$DATE) == 3, 1, 0)
db$Apr <- ifelse(month(db$DATE) == 4, 1, 0)
db$May <- ifelse(month(db$DATE) == 5, 1, 0)
db$Jun <- ifelse(month(db$DATE) == 6, 1, 0)
db$Jul <- ifelse(month(db$DATE) == 7, 1, 0)
db$Aug <- ifelse(month(db$DATE) == 8, 1, 0)
db$Sep <- ifelse(month(db$DATE) == 9, 1, 0)
db$Oct <- ifelse(month(db$DATE) == 10, 1, 0)
db$Nov <- ifelse(month(db$DATE) == 11, 1, 0)
```

## 2(a)
```{r}
lmObj <- lm(
    HOUSTNSA ~ Jan + Feb + Mar + Apr +
        May + Jun + Jul + Aug + Sep + Oct + Nov,
    data = db
)
# summary(lmObj)
```
<!-- Analysis -->
\begin {dmath}
    \widehat{HOUSTNSA} =      \underset{(4.196)}{92.871}
                        \ - \ \underset{(5.911)}{4.592}\ Jan
                        \ - \ \underset{(5.911)}{1.935}\ Feb
                        \ + \ \underset{(5.934)}{26.184}\ Mar
                        \ + \ \underset{(5.934)}{41.452}\ Apr
                        \ + \ \underset{(5.934)}{46.786}\ May
                        \ + \ \underset{(5.934)}{46.263}\ Jun
                        \ + \ \underset{(5.934)}{40.937}\ Jul
                        \ + \ \underset{(5.934)}{38.714}\ Aug
                        \ + \ \underset{(5.934)}{32.252}\ Sep
                        \ + \ \underset{(5.934)}{36.170}\ Oct
                        \ + \ \underset{(5.934)}{15.600}\ Nov
\end {dmath}


## 2(b)
```{r}
db2 <- read.csv2(
    file = "Assignment 2/Data/HOUSTNSA.csv",
    sep = ","
)

db2 <- db2 %>%
    as_tibble() %>%
    mutate(
        DATE = yearmonth(DATE),
        HOUSTNSA = as.numeric(HOUSTNSA)
    ) %>%
    as_tsibble()

db2$Feb <- ifelse(month(db2$DATE) == 2, 1, 0)
db2$Mar <- ifelse(month(db2$DATE) == 3, 1, 0)
db2$Apr <- ifelse(month(db2$DATE) == 4, 1, 0)
db2$May <- ifelse(month(db2$DATE) == 5, 1, 0)
db2$Jun <- ifelse(month(db2$DATE) == 6, 1, 0)
db2$Jul <- ifelse(month(db2$DATE) == 7, 1, 0)
db2$Aug <- ifelse(month(db2$DATE) == 8, 1, 0)
db2$Sep <- ifelse(month(db2$DATE) == 9, 1, 0)
db2$Oct <- ifelse(month(db2$DATE) == 10, 1, 0)
db2$Nov <- ifelse(month(db2$DATE) == 11, 1, 0)
db2$Dec <- ifelse(month(db2$DATE) == 12, 1, 0)

lmObj2 <- lm(
    HOUSTNSA ~ Feb + Mar + Apr + May + Jun +
        Jul + Aug + Sep + Oct + Nov + Dec,
    data = db2
)
# summary(lmObj2)
```
<!-- Analysis -->
### Steps
In order to formulate the linear regression, first we need to determine the intercept:
From equation 1 we can determine the values of each month because of the dummy variables.
$92.871 - 4.592  = c \rightarrow c = 88.280$, where the LHS is the month of Jan from calculated from equation 1. 

Next we need to determine the $\beta$ values for Feb - Dec. Since we know the intercept for the new equation, we can substitute it in.
\begin {align*}
92.871 + 1.935  &= 88.280 + \beta_2 \ Feb  \\ 
                &\rightarrow \ \beta_2 = 2.656 \\
                \\
92.871 + 26.184 &= 88.280 + \beta_3 \ Mar  \\ 
                &\rightarrow \ \beta_3 = 30.776 \\
                \\
92.871 + 41.452 &= 88.280 + \beta_4 \ Apr  \\ 
                &\rightarrow \ \beta_4 = 46.044 \\
                \\
92.871 + 46.786 &= 88.280 + \beta_5 \ May  \\ 
                &\rightarrow \ \beta_5 = 51.377 \\
                \\
92.871 + 46.263 &= 88.280 + \beta_6 \ Jun  \\ 
                &\rightarrow \ \beta_6 = 50.855 \\
                \\
92.871 + 40.937 &= 88.280 + \beta_7 \ Jul  \\ 
                &\rightarrow \ \beta_7 = 45.528 \\
                \\
92.871 + 38.714 &= 88.280 + \beta_8 \ Aug  \\ 
                &\rightarrow \ \beta_8 = 43.306 \\
                \\
92.871 + 32.252 &= 88.280 + \beta_9 \ Sep  \\ 
                &\rightarrow \ \beta_9 = 36.844 \\
                \\
92.871 + 36.170 &= 88.280 + \beta_10 \ Oct \\ 
                &\rightarrow \ \beta_{10} = 40.762 \\
                \\
92.871 + 15.600 &= 88.280 + \beta_11 \ Nov \\ 
                &\rightarrow \ \beta_{11} = 20.192 \\
                \\
92.871          &= 88.280 + \beta_12 \ Dec \\ 
                &\rightarrow \ \beta_{12} = 4.592 \\
                \\
\end {align*}

\begin {dmath}
    \widehat{HOUSTNSA} =      88.280
                        \ + \ 2.656 \ Feb
                        \ + \ 30.776\ Mar
                        \ + \ 46.044\ Apr
                        \ + \ 51.377\ May
                        \ + \ 50.855\ Jun
                        \ + \ 45.528\ Jul
                        \ + \ 43.306\ Aug
                        \ + \ 36.844\ Sep
                        \ + \ 40.762\ Oct
                        \ + \ 20.192\ Nov
                        \ + \ 4.592 \ Dec
\end {dmath}

## 2(c)
```{r}

```
<!-- Analysis -->

## 2(d)
```{r}

```
<!-- Analysis -->

## 2(e)
```{r}

```
<!-- Analysis -->
